---
title: "Essential R"
author: "Tinniam V Ganesh"
date: "April 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction
## About R
## Installing R and R Studio

##R basic
R is a very friendly lanuguage and flexible in many ways. While programming purists could complain 
that R is too weakly types, but I think this lack of rigitidy allows one to be productive.

So we can easily write statements without defining or declaring variables and their types. We can start
anywhere in the line, and continue onto the next line, There is no continuation character as in other
langiuages, no statement delimiters. As long as you end a line with an operator you can continue 
on to the next line. Also in R, the assignment is the "<-". Also when we write the variable on a 
separate line R displays the value of the element as can be seen in the 2nd line below

###Variable types
```{r}
c <- 5.26  # No variable definitions or declarations
c    # Display the value of c
a <- 5 * 3 +    # No line continuation character, no statement delimiters
     2/6
a
      c <- 4 + 2.679 +    # Start anywhere and continue to the next line by ending the line with an operator
                 23*7/8
```

#Note: In R we can display the value of a variable, matrix, dataframe by typing it on a separate
line in the R console as below
```{r}
c
```
The main vaiable types are 

a) Character
b) numeric

You can check the variable type by using the "class" command. This is a very useful command
and often times we may need to check the variable type before trying to use it in computaion/plots etc.
```{r}
x<- "Hello world" # Can also use x= "Hello World"
x
class(x)
y<-2.5
y
class(y)

```

###Vectors
An important R type is the vector. In a vector all elements are of the same type. If for example
you add a character variable to a numeric vector myvector2 as shown below, then myvector2 will become
a character vector
```{r}
#2.  Vectors - All elements of same type
myvector <- c(1,2,4,5,8,10,12)
myvector
myvector <- 2:12
myvector
charVector <- c("one","two","three") # character vector
charVector
myvector2 <- c(5,7,9,12)
class(myvector2)
myvector2 <-c(myvector2,"the")
class(myvector2)
class(myvector2[1])
```

You can get the length of the vector with
```{r}
length(myvector)
```
###Sequences
Vectors can also be created by using the "seq" command. The utility of sequences may not be immediately
apprarent, nevertheless they are very important and you will be forced to create and use
sequences in your code.
```{r}
# Creating vectors with sequences. In the example below, the sequence starts at 5 and ends at 50 with
# an increment of 5
myvector <- seq(5,50,by=5)
myvector

# Create a sequence between 6 & 43 with a period length of 7
seq(6,43,by=7)

# This is useful when we want 7 equal intervals between 6 & 43
seq(6,43,length=7)

```


###Repeating sequences
Repeating sequences can be created b
```{r}
a <- rep(5:9,times=3)
a
b <- rep(3:7,len=2)
b
```
# Q a<- 5.4
# What is the class of a?
# Let a <- seq(2:7,by=3)
# Let b <- c(a,"boy")
#Q What is the class of a & b
#Q Create a sequence from 12 to 200 and increment by 8
#Q Create a sequence from 8 to 51 and create 9 equally space numbers between them
#Q Create a repeating sequence from 3 to 7 and repeat it thrice


One of the nice things about R is that you can get help on R functions while coding without
going to Google, as long as you have installed the package in your environment. You can access
help on any specific topic/R function by typing ?<function> or help(<function>)
```{r}
# Get help on any topic using '?'
?seq

# or
help(seq)

```

A useful command is 'which' which returns the index where a value(s) are found
```{r}

myvector <- c(1,2,4,5,8,4, 10,12)
which(myvector ==4)
```

# Q a<- 5.4
# What is the class of a?
# Let a <- seq(2:7,by=3)
# Let b <- c(a,"boy")
#Q What is the class of a & b
#Q Create a sequence from 12 to 200 and increment by 8
#Q Create a sequence from 8 to 51 and create 9 equally space numbers between them
#Q Create a repeating sequence from 3 to 7 and repeat it thrice
#Q Check the help of 'rep' comand. What does the 'len' parameter indicate?

#### Matrix,lists and dataframes

Matrices
Matrices can be created as shown below with 5 rows and 4 columns
```{r}
# generates 5 x 4 numeric matrix
y<-matrix(1:20, nrow=5,ncol=4)
class(y)
```

Lists
In lists the elements can be dissimilar. The list belowhas the following objects
a) Character variable b) character vector c) matrix and 4) numeric varaible
```{r}
w <- list(name="Fred", mynumbers=charVector, mymatrix=y, age=5.3)
w
```

##5. Data frames
A data frame is the most common data types in R. Most data analysis is done on dataframes.
Dataframes can be thought of as Excel sheets with rows  and columns. Each column can be of
a different type but all elements in the column must be of the same type.
```{r}
d <- c(1,2,3,4)
e <- c("red", "white", "red", NA)
f <- c(TRUE,TRUE,TRUE,FALSE)
mydataframe <- data.frame(d,e,f)
names(mydataframe) <- c("ID","Color","Passed") # variable names
mydataframe
class(mydataframe)
```

###Subsetting
One of the most important operations that you will most often use in R is subsetting i.e.
we need to get a small subset of a vector, list, matrix or a dataframe. It is very important to
understand how to do this.
1. Subsetting vectors
```{r}
myvector <- seq(2,20,by=2)
myvector
```

Get the 4 element in my vector. Note in R the indices start from 1 to N and 0 to N-1 as is
common in C,C++,Java or Python.
```{r}
myvector[4]
```

Get a range of elements, specifically get all elements from the 3rd to 7th
```{r}
s <- myvector[3:7]
s
```
Subset based on a condition
Get all elements of myvector which are greater than 10
Here the fist statement creates a logical vector 'b' which is series of True and False. The logical
vector 'b' is used to subset myvector in the 3rd line to get the elements that are >10> The logical
vector 'b' is used as index to 'a' to get all elements > 10.

```{r}
b <- myvector>10
b
a <- myvector[b]
a
```

It is possible to Combine both in a single statement as shown below
```{r}
a <- myvector[ myvector>10]
```


Sometimes we may need to leave out a particular elements of a vector for e.g we would like
all elements except the 2nd element. In this case we can use the '-index' as below
```{r}
myvector[-2]
```

Subsetting a character vector. Here we subset based on a substring. In the code below
m is a character vector. We subset 'm' using the grep funvtion. This creates 'a' which is a
logical vector. We can then use this logical vector as an index to subset all elements
whicc contain the string "ab"
```{R}
m <- c("about","aboard","board","bus","cat","abandon")
a <- grep("ab",m)
a
# String with "ab"
containsAB <- m[a]
#Or
containsAB <- m[grep("ab",m)]
containsAB
```

###Operation on vectors
We can do arithmetic and logical operation on vectors if the vectors are of the same length

```{r}
myvector <- seq(2,20,by=2)
myvector
# Arithmetic operations on vectors
myvector1 <- seq(3,30,by=3)
myvector1

# Addition of 2 vectors
a <- myvector + myvector1
a
# Product
b <- myvector * myvector1
b
# Raising element to a power
myvector^3

#Logical operations
a <- myvector < myvector1
a

```

You can also do arithmetic and logical operations on vectors as long as the length of one of
the vectors is a multiple of the other

```{r}
a <- c(1,2,9,4,12,6)
b <- c(7,8)

a+b
a-b
a<b

```


###2.  Accessing elements in matrix
We can access elements in a matrix by specifying the row and the column index. We can also specify
the row and the column vector
```{r}
y<-matrix(1:20, nrow=5,ncol=4)
y
y[2,3]
```

###Matrix operations
Included below are some key matrix operations. R has several packages that deal with matrices
for various analysis. Some of the main operations are shown below

```{r}
A <- matrix(1:9,nrow=3,ncol=3)
B <- matrix(seq(5,10,length=9),nrow=3,ncol=3)
```

### Some key matrix operations

```{r}
# Matrix addition
C <- A+B
C
# Matrix subtraction
C <- A -B
C
#Element wise matrix multiplication
C = A*B                   
C
#Matrix multiplication
C = A %*% B 
C
```


###3. Subsetting lists
Lists can be subsetted using an index and a souble square bracket
```{r}
w <- list(name="Fred", mynumbers=charVector, mymatrix=y, age=5.3)
w


# To access an object in a list , use [[]]e.g
w[[2]]
# You can further access the elements in the object depending on the type
a = w[[2]] #a is a vector
a[2]
b=w[[3]]
b
#
b[2,3]
```

###4. Subsetting dataframes
To know how to subset data frames is extremely important because by and large you will be working
on subsetting dataframes. So do try to understand how to do this.R comes with already available 
data frames with which you can do any data science with data()
In the example below the in-built dataframe mtcars will be used for showing to subset dataframes

```{r}
head(mtcars,3)
# We can subset like matrices
# Dataframe 'df' are subsetted as
#df[row,column] 
mtcars[2,3]

# We can also subset by providing the row & column vectors as  df[row vector,column vector]
mtcars[2:5,3:6]


# If the row/row vector is not included then it means all rows
#Get all rows but only columns 3 to 7
mtcars[,3:7]


# If the column/column vector is not included then it means all rows
# Get all columns for rows 2 to 5
mtcars[2:5,]
```

To check the column names of a dataframe we can either use names() or colnames()
```{r}
names(mtcars)
colnames(mtcars)
```

To access a specific column by name use the <dataframe>$<columnname>
```{r}
mtcars$mpg
```

We can also subset based on a specific condition. To subset all rows of dataframe where mpg >15.0
we can use
# Find all rows where mpg >15
```{r}
a <- mtcars$mpg>15
a

#Subset the dataframe using this logical vector for the row vector. Leave
# the column empty
b <- mtcars[a,]
b
```

To subset on a compound condition we Can also write and assign the output to a logical vector 
a. We can then subset all rows based on this logical vector.
```{r}
a <- mtcars$mpg>15 & mtcars$mpg <25
a

# Subset using the logical vector 'a' and select all columns (no column vector)
c <- mtcars[a,]
c

```
#################################################################################
#Q5. Subset all rows for which the disp is >200
#Q6. Find all cars which are 6 cylinder 'cyl'
#################################################################################

# There are other ways of handling dataframes which will be shown below
#####################################################################################
###C. Some common directory and file handling commands
Get the current working directory
```{r}
getwd()
```

#Set the working directory. Note the forward slash
```{r}
setwd("C:/software/R")
getwd()
```

# Go up one directory and then the current directory
```{r}
setwd("..")

#Check this directory
getwd()
```

# Show all objects in the environment
```{r}
ls()
```
# Remove  a specific object for e.g. 'myvector which was created earlier
```{r}
rm("myvector")
```

#Remove all objects. This is a particularly useful command. We need this so that we can start
# with a clean slate and don't have variables left by some previous operations
```{r}
rm(list=ls())
```

#Show contents of dir
```{r}
dir()
```

# Other useful commands
# Create a directory
```{r}
dir.create("testdir")  
```
# Create a file
```{r}
file.create("myTest.R")  
```

#Check if file exists
```{r}
file.exists("mytest.R") 
```

# Rename a file
```{r}
file.rename("mytest.R","mytest2.R")     
```
# To print an object
```{r}
myvector <- seq(2,20,by=2)
myvector
myvector
print(myvector)
```

# For loops
There are 2 ways to write for loops. However one should 'for' loops if possible as 'for' loops
are very performance intensive. try to use vectors instead of 'for' loops wherever possible
```{r}
for(i in 1:10){
  print(i*5)
}

for(i in seq_along(myvector)){
  print(myvector[i])
}
```

# In R avoid for loops if possible. Try to use the while vector using sapply,lapply,tapply
# instead of for loops. for loops are very performance intensive to vector operations


### R functions. To create a function
# This is how a function is defined. The return value is usually made the last statement of the function. In the function below the return is the value 'c'
```{r}
product <- function(a, b){
  c <- a*b
  c
}
```
#Invoke the function
```{r}
product(7,12)
product(13,18)
```

##############################################################################
#Q.7. Create a function that raises a number to the 5th power
#Q.8. Create a function that takes mtcar and cyl as input and display all rows
# with this combination
#############################################################################

##########################################################################################

### Accessing elements in a dataframe
# There are several datasets available that comes with R
# The Dataframe is a tabular form of data with many rows and columns. The
# type of data can vary from column to column but are the same in each column
```{r}
data()
class(iris)
```

# Create a new data frame from iris. Iris is a small plant.
# See https://en.wikipedia.org/wiki/Iris_(plant)
# Display the iris dataframe by typing it on a separate line
```{r}
iris
```


# Check the size of the data. This is a very important command and you will be using it very often
```{r}
dim(iris)
```

#Check the 1st few rows of the data frame
```{r}
head(iris)
```

A very useful command to check the type of a dataframe is. The 'str' command lists the
columns in the dataframe, their type and also displays a few representative values from each
column
```{r}
str(iris)
```

We could also use the 'summary' command to get an overall view of dataframe.However the 'str' command
is more powerful than the 'summary' command

```{r}
summary(iris)
```

To check the column names of iris we cause one of the following. They are both same

```{r}
colnames(iris) 

#or 
names(iris)
```

We often need to check the class of every column. We could do this one by one or simply use
```{r}
sapply(iris,class)
```

To inspect the data. Display top 6 and bottom 6 of the dataframe
```{r}
head(iris,6)
tail(iris,6)
```


###Subset data
1. Display 1-6 rows and 2-4 columns. This similar to how we subsetted the matrix
We can use df[row,column] or df[row vector, column vector]
```{r}
iris[1:6,2:4]
```

2. Display all rows  for columns 1:3. If the row/row vector is omitted then display all rows
```{r}
iris[,1:3]
```

3. Display all columns for rows 1:5. If the column/column vector is omitted then display all columns
```{r}
iris[1:5,]
```

##############################################################################
#Q9.  Display all rows and columns 2,3,5 of iris
#Q10. Display all rows and columns 2:4 of iris
##############################################################################
#Column names
As before we can check the column names 
```{r}
colnames(iris)
```

To rename columns,  we use the same function 'colnames' but we assign to it a character vector as shown
```{r}
colnames(iris) <- c("lengthOfSepal","widthOfSepal","lengthOfPetal","widthOfPetal","Species")
colnames(iris)
```

Now if we view the dataframe the column names would have been changed
```{r}
View(iris)
```

Here are a few other ways of subsetting the data. We can directly refer to a column in a dataframe by using
the df$<column name> to refer to an entire column
```{r}
iris$Species
```

Now that we can refer to a specific column in a dataframe we can subset on specific values in the column
For e.g. we can subset all rows which have "setosa" as the Species. Below we use '==' sign to check
if the species is 'setosa' and we asign the output of this logical operation to a logical vector 'a'. If we display the value of 'a' we see that it is a logical vector containing TRUE & FALSE
```{r}
a <- iris$Species == "setosa"
a
```

If we display the value of 'a' we see that it is a logical vector containing TRUE & FALSE
```{r}
a
```

Subset these rows. The "," indicates include all colums for these rows. Here for the row vector we use the
logical vector and we leave the column vector indicating we will require all columns
```{r}
b <- iris[a,]
```

Check the dimemsion ofthe 'b' dataframe
```{r}
dim(b)
b
```

Using other functions on a column We can refer to the column of a dataframe by using '$' sign
```{r}
colnames(b)
b$widthOfSepal
```

Compute the mean and standard deviation of the Width of the sepal as below

```{r}
meanSepal <- mean(b$widthOfSepal)
meanSepal
sdSepal <- sd(b$widthOfSepal)
sdSepal
```


A very useful function is 'sapply' with which can apply  a function across vectors. In the R statement below
we have 4 column vectors of iris and we apply the function 'mean' to each  of these column vectors
```{r}
m <- sapply(iris[1:4],mean)
m
```

To  take a quick look at a specific column the box plot is particulary useful. To take a look at the 
documentation for a boxplot type ?boxplot
```{r}
?boxplot
```

Create a bixplot for the length of the sepal column of the Iris dataframe. The title of the plot requires the
use of the 'main=' argument
```{r}
boxplot(b$lengthOfSepal,main="Length of Sepal")
```


Often we may need to display a boxplot for all the 4 numerical column vectors of Iris namely - widthOfSepal, lengthOfSepal, widthOfPetal & lengthOfPetal. The way to do this is to use the 'par' function with the 'mfrow' which specifies that we would like  4 plots in a 2 x 2 matrix setup.Draw all 4 plots. The 'mar=' parameter indicates the margins
```{r}
par(mfrow=c(2,2)) # Set 2 rows x 2 columns
par(mar=c(4,4,2,2)) # Set the margins
boxplot(b$lengthOfSepal,main="Length of Sepal")
boxplot(b$widthOfSepal,main="Width of Sepal")
boxplot(b$lengthOfPetal,main="Length of Petal")
boxplot(b$widthOfPetal,main="Width of Petal")
```

Reset the display with the following call. Otherwise you will get a 2 x 2 plots
```{r}
dev.off()
```

A useful function to to check how data looks is the pairs function.This function will take pairs of columns and do a scatter plot. This will tell us how the columns are correlated
```{r}
pairs(iris)
```

R also provides a way to compute the correlation numerically as follows
#Length of Petal is positively correlated to width of Petal and length of Sepal
```{r}
cor(iris[,1:4])
```

#Q6, Draw a box plot of mpg and displ from mtcars
```{r}
head(mtcars,5)
par(mfrow=c(1,2)) # Set 2 rows x 2 columns
par(mar=c(4,4,2,2)) # Set the margins
boxplot(mtcars$mpg,main="Miles per gallon")
boxplot(mtcars$disp,main="Displacement")
dev.off()
```

Make a regular scatter plot using the R base plotting function plot(x,y)
```{r}
plot(iris$widthOfSepal,iris$lengthOfSepal)
```

To this plot add the x-axis & y-axis labels along with the title of the plot use xlab, ylab and main parameters


```{r}
plot(iris$widthOfSepal,iris$lengthOfSepal,
     main="Iris - Length of Sepal vs Width of Sepal", xlab="Sepal Width",
     ylab="Sepal length")
```
########################################Cleaning data ##############################################

The problem with the IRIS data set is that it is neat and tidy. In fact most of the datasets that are
available on R through the function data() are clean. However almost all real world data is unclean, has many missing values, spaces for values or othr junk characters. We must first  'clean' the data before applying any visualization or Machine Learning models on the data. So I would like to present a small sample data
which has many of the issues that we will encounter in real data. 

The following call to read.csv() does not convert strings in the data frame as factors which is
specified in the stringAsFactors=FALSE. Further the argument 'na.strings' specified which characters need to
be considered as NA for e.g. we could specific na.strings=c(NA,"-"," ",""). This means that all values
in the data frame NA, "-", " " and "" will be substituted with NA in the dataframe that is read into memory.
For more details check the help of read.csv 
Check the help of read.csv using ?read.csv

# Read the tendulkar dataframe
```{r}
tendulkar= read.csv("tendulkar.csv",stringsAsFactors = FALSE,na.strings=c(NA,"-"))
```

The 'summary' is a useful command for describing the overall content and structure of a dataframe
```{r}
summary(tendulkar)
```

The 'str' command is another command to get an overall description of the dataframe. The 'str' command lists
all the columns, the column type and also displays a representative sample of the column values. I find the 'str' command more useful than the 'summary' command
```[r]
str(tendulkar)
```

Check the names of the columns for the dataframe tendulkar
```{r}
colnames(tendulkar)
```

Also check the dimensions of the tendulkar dataframe. It can be seen that the dataframe is 
```{}
dim(tendulkar)
```

#Display top 5 and bottom 5 rows
```{r}
head(tendulkar)
tail(tendulkar)
```

As mentioned before if we want to check the class of all columns we can use the sapply() function to
get the types of all columns
```{r}
sapply(tendulkar,class)
```


# View the data frame
To view a dataframe in RStudio you can use the View(df)
```{r}
View(tendulkar)
```

# Real world data will require a lot of cleaning before you can use it

The following list of commands are used to clean the tendulkar dataframe. In real world data there will be  a need to clean up data in a variety of ways depending on how unclean the data is. There could some columns or
rows with completly or partially missing values. There could be elements where there are garbage characters and so on. The commands that will be needed will vary but at the end of the cleaning process, all columns should be of the same type. Here are the commands to clean tendulkar dataframe

a) Remove all rows which have DNB (Did Not Bat) in the 'Runs' column, or in other words select rows which do not have the 'DNB' value. This is given by the '!=' or not 'equal to' logical operator. Assign the result to a logical vector a
```{r}
a <- tendulkar$Runs != "DNB"
a
```
Now subset all rows based on the logical vector 'a', leave out the column vector while subsetting indicating that all columns must be selected.
```{r}
tendulkar <- tendulkar[a,]
dim(tendulkar)
```

b) Next remove all rows with TDNB (Team Did Not bat) and assign the result to a logical vector 'c'. Use 'c' to subset the tendulkar dataframe

```{r}
c <- tendulkar$Runs != "TDNB"
c
tendulkar <- tendulkar[c,]
```

c) Further remove all rows which have 'absent' in the Runs column and subset as before
```{r}
d <- tendulkar$Runs != "absent"
tendulkar <- tendulkar[d,]
```

Check the size of the tendulkar dataframe
```{r}
dim(tendulkar)
```

d) Finally remove the "*"" indicating 'not out' 
```{r}
tendulkar$Runs <- gsub("\\*","",tendulkar$Runs)
```

Check the size of the cleaned tendulkar dataframe
```{r}
dim(tendulkar)
```

Another useful function for cleaning data is complete.cases. Check the documentation on
'complete.cases()' 
```{r}
?complete.cases
```

e) Check if all rows are complete and have values in all columns and assign result to logical vector 'c'. Use
this logical vector to subset the dataframe to remove incomplete rows
```{r}
c <- complete.cases(tendulkar)
c
#Subset the rows which are complete
tendulkar <- tendulkar[c,]
```

Check the size of the tendulkar dataframe after all the above steps (a) to (e) for cleaning
```{r}
dim(tendulkar)
```

Now the data is clean. You can use View(tendulkar) in RStudio to ensure that data is indeed clean
```{r}
head(tendulkar,10)
```


########################################Plotting#############################################
### Base plot
# Do some base plotting
plot(tendulkar$BF,tendulkar$Runs)
#Add title and x and y labels
plot(tendulkar$BF, tendulkar$Runs,pch=18, main="Tendulkar Runs scored vs Balls Faced",
     xlab="Balls Faced", ylab='Runs')

?lm

# Compute and  plot a linear regression lime
# Here Runs = a0 + a1 * BF
fit <-lm(tendulkar$Runs ~tendulkar$BF)
fit
summary(fit)
abline(fit,lty=5,lwd=3,col="blue")


#Fitting a 2nd order polynomial
#Fit a 2nd order polynomial
# Here Runs = a0 + a1 * BF + a2 * BF^2
fit <- lm(tendulkar$Runs~poly(tendulkar$BF,2,raw=TRUE))

xx <- seq(from=0,to = max(tendulkar$BF),by=20)
yy <- NULL
for (i in seq_along(xx)) {
    yy[i] <- fit$coefficients[3] * xx[i]^2 + fit$coefficients[2] * xx[i] + fit$coefficients[1] 
    
}

lines(xx,yy,col="red",lwd=2.0)




# Q7, Plot the distribution of Minutes in crease and Runs Scored
# Q8 Fit a linear regression line
# Q9. Create a 2nd order polynomial between Runs scored and 4s
# Q10. Create a multivariate regression of Runs scored vs Balls Faced + Mins at Crease
#Here
# Runs = a0 + a1 * BF + a2 * Mins
fit <- lm(tendulkar$Runs ~tendulkar$BF + tendulkar$Min)
fit


#Plot the histogram of runs for Tendulkar
df = tendulkar
hist(df$Runs, main="Tendulkar's frequency of runs vs  run ranges",
     xlab="Runs")

abline(v=median(df$Runs),col="blue",lwd=3.0)
#Q11 Add a red line showing the mean runs for Tendulkar
abline(v=mean(df$Runs),col="red",lwd=3.0)

# A look package dplyr. One of the most useful package for manipulating data in
# data frames
#install.packages("dplyr")

#Call the library to the package to include dplyr
library(dplyr)

# Select columns
colnames(tendulkar)

# Key functions select, filter, arrange, pipe,mutate, summarise

# Subset specific columns by name of column. Note there is no $ sign or quotes to refer to
# columns
# Note there is no need for the $ to reference column names.
df1 <- select(tendulkar, Runs,Mins,BF)
head(df1)

# This can be also written as below, where %>% is known as the 'pipe'
df1 <- tendulkar %>% select(Runs,Mins,BF)
df1

#Subset rows where the RUns are > 50 and less than 101
df2 <- filter(tendulkar,Runs>50 & Runs < 101)
head(df2,5)
# Also as
# df2 <- tendulkar %>% filter(Runs>50 & Runs < 101)

# Using dplyr to clean the Tendulkar dataframe
tendulkar= read.csv("tendulkar.csv",stringsAsFactors = FALSE,na.strings=c(NA,"-"))
dim(tendulkar)
tendulkar <- tendulkar %>% filter(Runs != "DNB")
dim(tendulkar)

#Q12 
tendulkar <-  tendulkar %>% filter(Runs != "TDNB")
dim(tendulkar)

#Q13 Remove rows for which Tendulkar was "absent"
tendulkar <- tendulkar %>% filter(Runs != "absent")
#Finally we remove '*' as before
tendulkar$Runs <- gsub("\\*","",tendulkar$Runs)
class(tendulkar$Runs)
tendulkar$Runs <- as.numeric(tendulkar$Runs)

# Get only complete cases as before
c <- complete.cases(tendulkar)
tendulkar <- tendulkar[c,]
dim(tendulkar)

# There are more interesting conditions with which you can filter and select rows
# columns. Check with ?select & ?filter

# Use the arrange function to arrange columns in descending order of Runs
descRuns <- arrange(tendulkar,desc(Runs))
class(tendulkar$Runs)
head(descRuns)

#Create a new column with Strike rate
tendulkar <- tendulkar %>% mutate(StrikeRate=(Runs/BF)*100)


# The Pipe function is extremely useful in dplyr
a <- group_by(tendulkar, Ground)
head(a)

# The following command in a single line does the follwoing
#1. Uses the data frame tendulkar (2nd parameter)
#2. Groups the runs scored by ground
#3. Computes the mean runs in each group
#4. Arranges in descending order

tendulkar1 <- tendulkar %>% group_by(Ground) %>% summarise(meanRuns= mean(Runs)) %>%
  arrange(desc(meanRuns))


# Plot the result as a barplot
barplot(tendulkar1$meanRuns,names=tendulkar1$Ground,las=2,ylab="Average Runs",
        col=rainbow(length(tendulkar1$meanRuns)),main="Tendulkar's average runs in Grounds",cex.names=0.8)


#### Grammar of Graphics
#### ggplot2
library(ggplot2)

p <- ggplot(tendulkar) + aes(x=Runs,y=BF) + geom_point()
p <- p + xlab("Runs") + ylab("Balls Faced") + ggtitle("Tendulkar - Runs vs Balls Faced")
p

ggplot(tendulkar,aes(x=Runs,y=BF))+ geom_point() +
  xlab("Runs") + ylab("Balls Faced") + ggtitle("Tendulkar - Runs vs Balls Faced")

# Fit a smoothed regression line
ggplot(tendulkar,aes(x=Runs,y=BF))+ geom_point() + geom_smooth(method="loess") +
  xlab("Runs") + ylab("Balls Faced") + ggtitle("Tendulkar - Runs vs Balls Faced") +
  theme(plot.title = element_text(size=16, face="bold",hjust=0.5))


# Using dplyr and ggplot2
tendulkar1 <- tendulkar %>% group_by(Opposition) %>% summarise(meanRuns= mean(Runs)) %>%
    arrange(desc(meanRuns))
head(tendulkar1,10)

ggplot(tendulkar1,aes(x=Opposition,y=meanRuns,fill=Opposition)) + geom_bar(stat="identity")

# Rework this to reorder from biggest to smallest and title and labels 
ggplot(tendulkar1,aes(x=reorder(Opposition,-meanRuns),y=meanRuns,fill=Opposition)) + 
    geom_bar(stat="identity") +
    ggtitle("Tendulkar's Mean Runs against opposition") +
    xlab("Opposition") + ylab("Mean Runs")

############################################################################
#Q13. Create a barplot of Tendulkar's mean runs inm different grounds

#Q14 Compute and list the mean Runs and standard deviation  against opposition
################################################################################

## Going back to the Iris example
names(iris)
# Rename columns, if needed, to something more is easy to refer to
colnames(iris) <- c("lengthOfSepal","widthOfSepal","lengthOfPetal","widthOfPetal","Species")
colnames(iris)

setosa <- iris %>% filter(Species == "setosa")
#Correlation between columns
pairs(setosa)
a <- setosa %>% select(lengthOfSepal,widthOfSepal)
plot(a$lengthOfSepal,a$widthOfSepal,xlab="Length of Sepal",ylab="Width of Sepal",
     main="Length vs Width of Sepal of Setosa")
l <-lm(a$widthOfSepal~a$lengthOfSepal)
abline(l,lty=5,lwd=3,col="blue")

#
str(iris)
# This can be with ggplot as follows for all species
ggplot(iris,aes(x=lengthOfSepal,y=widthOfSepal,colours=Species)) + geom_point() +
  geom_smooth(method="loess") + facet_wrap(~Species) +
  xlab("Length of Sepal") + ylab("Width of Sepal")


# Package lubridate.
# This package is useful for handling date files in most formats
#dmy,ymd,dmyhhmmss,dd-mm-yy hh:mm:ss
#install.packages9"lubridate")
library(lubridate)
start<-ymd("20110604")
day(start)
month(start)
year(start)

end=mdy("08-04-2011")
timespan <- end-start
timespan
c <- end+5
c

# There is a  lot more that can be done with lubridate
head(tendulkar,5)

# Creating a moving average of runs for Tendulkar with ggplot
runs <- tendulkar %>% select(Runs,Start.Date)
runs$Start.Date <- dmy(runs$Start.Date)
head(runs)
ggplot(runs,aes(x=Start.Date,y=Runs)) + geom_line(colour="darkgrey") +
  geom_smooth(method="loess") + ggtitle("Tendulkar's moving average of runs") +
  theme(plot.title = element_text(size=16, face="bold",hjust=0.5))


a <- PlantGrowth %>% select(weight,group) %>% group_by(group) %>%
    summarize(meanWeight=mean(weight),stdWeight=sd(weight))
    

ggplot(a,aes(x=group,y=meanWeight,fill=group)) + geom_bar(stat="identity") +
    ggtitle("Group vs Mean Weight with Error bar") + xlab("Group") + ylab("Mean Weight") +
    geom_errorbar(width=.1, aes(ymin=meanWeight-stdWeight, 
                                ymax=meanWeight+stdWeight))


##################################################################################################
######################################################################################################
##############################################################################
#Q.7. Create a function that raises a number to the 5th power
power5 <- function(n){
     a <-n^5
     a
}
#Q8. Create a function that takes mtcar and cyl as input and display all rows
# with this combination

carsCyl <- function(df,numcyl){
  a <- df$cyl == numcyl
  mtcars[a,]
}

##############################################################################
#Q9.  Display all rows and columns 2,3,5 of iris
cols <- c(2,3,5)
iris[,cols]
#Q10. Display all rows and columns 2:4 of iris
iris[,2:4]
##############################################################################

#######################################################################
##############################################################################
#Q13. Create a barplot of Tendulkar's mean runs inm different grounds
tendulkar1 <- tendulkar %>% group_by(Ground) %>% summarise(meanRuns= mean(Runs)) %>%
  arrange(desc(meanRuns))
top10 <- head(tendulkar1,10)
ggplot(top10,aes(x=reorder(Ground,-meanRuns),y=meanRuns,fill=Ground)) + 
  geom_bar(stat="identity") +
  ggtitle("Tendulkar's Mean Runs in Grounds") +
  xlab("Grounds") + ylab("Mean Runs")

#Q14 Compute and list the mean Runs and standard deviation  against opposition
tendulkar1 <- tendulkar %>% group_by(Opposition) %>% 
  summarise(meanRuns= mean(Runs),sdRuns=sd(Runs)) %>%
  arrange(desc(meanRuns))
top10 <- head(tendulkar1,10)
top10
ggplot(top10,aes(x=reorder(Opposition,-meanRuns),y=meanRuns,fill=Opposition)) + 
  geom_bar(stat="identity") +
  geom_errorbar(width=.1, aes(ymin=meanRuns-sdRuns, 
                              ymax=meanRuns+sdRuns)) +
  ggtitle("Tendulkar's Mean Runs against opposition") +
  xlab("Opposition") + ylab("Mean Runs")


############################################################################################
