---
title: "Essential R"
author: "Tinniam V Ganesh"
date: "April 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

##R basic
R is a very friendly lanuguage and flexible in many ways. While programming purists could complain 
that R is too weakly types, but I think this lack of rigitidy allows one to be productive.

So we can easily write statements without defining or declaring variables and their types. We can start
anywhere in the line, and continue onto the next line, There is no continuation character as in other
langiuages, no statement delimiters. As long as you end a line with an operator you can continue 
on to the next line. Also in R, the assignment is the "<-"
```{r}
c <- 5.26  # No variable definitions or declarations
a <- 5 * 3 +    # No line continuation character, no statement delimiters
     2/6
a
      c <- 4 + 2.679 +    # Start anywhere and continue to the next line by ending the line with an operator
                 23*7/8
c
```
The main vaiable types are 

a) Character
b) numeric

You can check the variable type by using the "class" command. This is a very useful command
and often times we may need to check the variable type before trying to use it in computaion/plots etc.
```{r}
x<- "Hello world" # Can also use x= "Hello World"
x
class(x)
y<-2.5
y
class(y)

```



